Unprepared colonoscopy identifies microbiota specific to the proximal and distal human colon

### 2017-07-21 submission/test.Rmd
found the bug, this compiles for now

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
knitr::opts_chunk$set("tidy" = TRUE)
knitr::opts_chunk$set("echo" = FALSE)
knitr::opts_chunk$set("eval" = TRUE)
knitr::opts_chunk$set("warning" = FALSE)
knitr::opts_chunk$set("cache" = FALSE)

@@

The human colon is an ecosystem comprised of several different microenvironments inhabited by resident bacterial members of the microbiome. Concentrations of oxygen, water and anti-microbial peptides change along the gut axis and influence what populations of microbes reside in each location. Microenvironments differ not only longitutinally along the colon, but latitudinally from the epithelium to mucosa to intestinal lumen, offering several sites for different microbial communities to flourish. The identity of these specific microbes and communities are important for understanding the etiology of complex colon diseases such as Inflammatory Bowel Disease (IBD) and Colorectal Cancer (CRC). IBD and CRC are known to be preceded or accelerated by perturbations in gut microbes (@Yamauchi2012, @Forbes2016, @Halfvarson2017). The severity, symptoms, morbidity and mortality of these diseases is known to vary based upon the biogeographical location in which they occur. For instance, CRC tumors that arise on the distal side of the colon are infiltrating lesions that present with painful symptoms (@Benedix2010). In contrast, 47% of CRCs are caused by proximal-sided colon tumors that are sessile and form along the wall of the colon, often remaining asymptomatic until advanced carcinogenesis (@Benedix2010). The distal and proximal sides of the colon differ in the amount of inflammation present and the genomic instability of precancerous cells, respectively, in addition to variation in the previously mentioned chemical gradients (@Yamauchi2012, @Albenberg2014, @Donaldson2015). In IBD patients, disease flares in the distal colon are usually indicative of ulcerative colitis (UC), whereas Crohn's disease (CD) patients typically experience disease in the small intestine, ileum and proximal colon (@Forbes2016). UC presents as large and highly inflammed mucosal ulcers, where as CD lesions are often smaller and have areas of normal tissue distributed amongst the flare (@Forbes2016). Thus, given the varied physiology of the proximal-distal axis of the colon and known differences in disease patterns at these sites, symbiotic microbes and their metabolites likely vary as well, and may influence the heterogenous disease prognoses of IBD and CRC. Because CRC is a long-term complication of IBD, the distribution of microbes is important to understanding the pathophysiology of both diseases.

Several recent findings have shown that development and progression of IBD or CRC can be attributed to specific molecular events as a result of interactions between the gut microbiota and human host (@Yamauchi2012, @Halfvarson2017). For instance, comparison of the bacteria present on CRC tumors with those found on nearby healthy tissue has identified specific species that are tumor-associated (@Dejea2014). These species include the oral pathogens _Fusobacterium nucleatum_ and _Porphyromonas asacharolytica_. Interestingly, these periodontal pathogens have been highly predictive of whether a patient had CRC tumors or not in our human stool classification studies (@Baxter2016). _F. nucleatum_ has also been found to be elevated in the stool and biopsies of patients with IBD as compared to healthy controls (Strauss2011). Furthermore, studies of _F. nucleatum_ isolated from mucosal biopsies showed that more invasive  _F. nucleatum_ positively correlates with IBD disease level (Strauss2011). Like many intestinal pathogens, the bacteria appear to have a high-impact despite being lowly-abundant in the community (@Forbes2016). The molecular capabilities of these rare taxa may contribute to the colonic disease state. These studies often examined only shed human stool or the small intestine, preventing fine-resolution analysis of paired samples from the proximal and distal sides of the colon. Similarly, comparisons of on- or off-tumor/lesion bacteria rarely have matched tissue from the other side of the colon from the same patient, limiting what conclusions can be drawn about the colonic microbiome overall, let alone at that specific site. Due to these limitations, the contribution of the gut microbiota to IBD and CRC disease location in the colon is largely undefined. Characterizing these communities could provide needed insight into disease etiology, including how the disruption of the healthy community could promote the initiation or proliferation of the distinct proximal and distal CRC tumors or IBD flares.

@@

#### Microbial membership and diversity of the proximal and distal colon 
Lumenal and mucosal samples were collected from the proximal and distal colon of 20 healthy humans that had not undergone bowel preparation (Figure 1). Participants also collected stool at home one week prior to the procedure. To characterize the bacterial communities present at these sites, 16S rRNA gene sequencing was performed on extracted DNA from each sample. Each site was primarily dominated by _Firmicutes_ and _Bacteriodetes_ (Figure 2A), consistent with known variability in human microbiome research (@LloydPrice2016). Likewise, samples had varying levels of diversity at each site, irrespective of the individual (Figure 2B). For example, the proximal mucosa was more diverse than the distal for some individuals while the opposite was true for others. Therefore we could not identify a clear pattern of changes in microbial diversity along the gut axis.

@@

#### Random Forest classification models identify important OTUs on each side
To identify OTUs that were distinct at each biogeographical site, we constructed several Random Forest models trained using OTU abundances. We built the first model to classify the lumen versus mucosal samples for the proximal and distal sides, independently (Figure 4A). The constructed model used ((Xopt)) features for the proximal and ((Xopt)) for the distal. The models performed well when classifying these samples (`r round(cv10f_roc_proximal_bs$auc,3)` and `r round(cv10f_roc_distal_bs$auc, 3)`, respectively). The OTUs that were most predictive of each site are identified by their greatest mean decrease in accuracy when removed from the model. For distinguishing the proximal lumen and mucosa, OTUs from the _Bacteriodes_, _Actinomyces_, _Psuedomonas_ and two OTUs from the _Enterobacteraceae_ genera were differentially abundant (Figure 4B). The model classifying the distal lumen and mucosa identified OTUs from _Turicibacter_, _Finegoldia_, _Peptoniphilus_ and two OTUs from the _Anaerococcus_ genera that could distinguish lumen from mucosal samples (Figure 4C). These results indicate that there are fine differences between the different sites of the colon, and that these can be traced down to specific OTUs on each side.

@@

>>>>

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=TRUE}
knitr::opts_chunk$set("tidy" = TRUE)
knitr::opts_chunk$set("echo" = FALSE)
knitr::opts_chunk$set("eval" = TRUE)
knitr::opts_chunk$set("warning" = FALSE)
knitr::opts_chunk$set("cache" = TRUE)

@@

The human colon is an ecosystem comprised of several different microenvironments inhabited by resident bacterial members of the microbiome. Concentrations of oxygen, water and anti-microbial peptides change along the gut axis and influence what populations of microbes reside in each location. Microenvironments differ not only longitutinally along the colon, but latitudinally from the epithelium to mucosa to intestinal lumen, offering several sites for different microbial communities to flourish.

*2 sentences in here are broken*

The severity, symptoms, morbidity and mortality of these diseases is known to vary based upon the biogeographical location in which they occur. For instance, CRC tumors that arise on the distal side of the colon are infiltrating lesions that present with painful symptoms (@Benedix2010). In contrast, 47% of CRCs are caused by proximal-sided colon tumors that are sessile and form along the wall of the colon, often remaining asymptomatic until advanced carcinogenesis (@Benedix2010). The distal and proximal sides of the colon differ in the amount of inflammation present and the genomic instability of precancerous cells, respectively, in addition to variation in the previously mentioned chemical gradients (@Yamauchi2012, @Albenberg2014, @Donaldson2015). In IBD patients, disease flares in the distal colon are usually indicative of ulcerative colitis (UC), whereas Crohn's disease (CD) patients typically experience disease in the small intestine, ileum and proximal colon (@Forbes2016). UC presents as large and highly inflammed mucosal ulcers, where as CD lesions are often smaller and have areas of normal tissue distributed amongst the flare (@Forbes2016). Thus, given the varied physiology of the proximal-distal axis of the colon and known differences in disease patterns at these sites, symbiotic microbes and their metabolites likely vary as well, and may influence the heterogenous disease prognoses of IBD and CRC. Because CRC is a long-term complication of IBD, the distribution of microbes is important to understanding the pathophysiology of both diseases.

Several recent findings have shown that development and progression of IBD or CRC can be attributed to specific molecular events as a result of interactions between the gut microbiota and human host (@Yamauchi2012).

_missing last half of second paragraph here, hasnt been fully tested_

@@

#### Microbial membership and diversity of the proximal and distal colon 
Lumenal and mucosal samples were collected from the proximal and distal colon of 20 healthy humans that had not undergone bowel preparation (Figure 1). Participants also collected stool at home one week prior to the procedure. To characterize the bacterial communities present at these sites, 16S rRNA gene sequencing was performed on extracted DNA from each sample. Each site was primarily dominated by _Firmicutes_ and _Bacteriodetes_ (Figure 2A), consistent with known variability in human microbiome research (cite). Likewise, samples had varying levels of diversity at each site, irrespective of the individual (Figure 2B). For example, the proximal mucosa was more diverse than the distal for some individuals while the opposite was true for others. Therefore we could not identify a clear pattern of changes in microbial diversity along the gut axis.

@@

#### Random Forest classification models identify important OTUs on each side
To identify OTUs that were distinct at each biogeographical site, we constructed several Random Forest models trained using OTU abundances. We built the first model to classify the lumen versus mucosal samples for the proximal and distal sides, independently (Figure 4A). The constructed model used ((Xopt)) features for the proximal and ((Xopt)) for the distal. The models performed well when classifying these samples (`r cv10f_roc_proximal_bs$auc` and `r cv10f_roc_distal_bs$auc`, respectively). The OTUs that were most predictive of each site are identified by their greatest mean decrease in accuracy when removed from the model. For distinguishing the proximal lumen and mucosa, OTUs from the _Bacteriodes_, _Actinomyces_, _Psuedomonas_ and two OTUs from the _Enterobacteraceae_ genera were differentially abundant (Figure 4B). The model classifying the distal lumen and mucosa identified OTUs from _Turicibacter_, _Finegoldia_, _Peptoniphilus_ and two OTUs from the _Anaerococcus_ genera that could distinguish lumen from mucosal samples (Figure 4C). These results indicate that there are fine differences between the different sites of the colon, and that these can be traced down to specific OTUs on each side.

@@

##this is file from commit right before broken commit

#testing intro
